---
title: "Capstone Project EDA"
author: "Michael An"
date: "April 7, 2016"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

 source("/media/mike/UNTITLED/Foundations_Course/Capstone/exploratory_script_load.R")
```

<!-- #Just load larger bin data -->
<!-- ```{r} -->
<!-- load("4width_binned_data.RData") -->
<!-- ``` -->

<!-- #Alternate dataset, time points binned into larger intervals -->
<!-- ```{r} -->

<!--  #pick out the biological active time points -->
<!--  activity.bio.pts <- subset(c1045.as.column, subset = BrainMask != 0, -->
<!--         select = c(FALSE,FALSE,FALSE,paradigm$Biological == 1)) %>% -->
<!--    t() %>% data.frame() -->


<!-- rest.bio.pts <- subset(c1045.as.column, subset = BrainMask != 0, -->
<!--        select = c(FALSE,FALSE,FALSE,paradigm$Random == 1)) %>% -->
<!--    t() %>% data.frame() -->

<!-- binwidth <- 4 -->
<!--  temp <- aggregate(activity.bio.pts, -->
<!--                    list(as.factor(binwidth*floor((1:nrow(activity.bio.pts))/binwidth))), -->
<!--                                        mean) -->

<!--  c1045.as.column <- cbind( -->
<!--    subset(c1045.as.column, subset = BrainMask != 0, -->
<!--         select = c(1,2,3)), -->
<!--         t(temp[,-1])) -->

<!-- paradigm <- data.frame(rep(1, nrow(temp))) -->
<!-- names(paradigm) <- "Biological" -->


<!-- rm(list = "temp") -->

<!--  temp <- aggregate(rest.bio.pts, -->
<!--                    list(as.factor(binwidth*floor((1:nrow(rest.bio.pts))/binwidth))), -->
<!--                                        mean) -->
<!--  c1045.as.column <- cbind(c1045.as.column, -->
<!--         t(temp[,-1])) -->

<!-- paradigm <- rbind(paradigm, data.frame(Biological = rep(0, nrow(temp)))) -->
<!-- paradigm$Random <- 1-paradigm$Biological -->
<!-- paradigm <- transform(paradigm, Biological = as.logical(Biological), Random = as.logical(Random)) -->

<!--  rm(list = "temp") -->


<!-- ``` -->


#Alternate Dataset: Fewer runs of data
```{r}
num.runs <- 4
temp <- subset(c1045.as.column, subset = BrainMask != 0)
#temp[,c(4:9,(num.runs*12+9):(ncol(c1045.as.column)))] <- list(NULL)

skip <- 1
#skip the first 2 runs, get the next "num.runs" runs
temp[,c(4:9,(10):(12*skip+9),((num.runs+skip)*12+9+1):(ncol(c1045.as.column)))] <- list(NULL)


#temp2 <- rep(1,nrow(paradigm))
#temp2[c(1:6,(num.runs*12+6):nrow(paradigm))] <- 0
#paradigm <- paradigm[-c(1:6,(num.runs*12+6):nrow(paradigm)),]

#skip the first 2 runs, get the next "num.runs" runs
paradigm <- paradigm[-c(1:6,7:(skip*12+6),((num.runs+skip)*12+6+1):nrow(paradigm)),]
c1045.as.column <- temp

```


##Histogram of time activity points for (L) biological motion time points, (M) random motion time points, (R) all time points
```{r}

#plot histograms of time activity points for different task, over all voxels
p1 <- ggplot(data=stack(subset(c1045.as.column, subset = BrainMask != 0), 
                select = c(FALSE,FALSE,FALSE,paradigm$Biological) )) +
 geom_histogram(aes(x=values))

p2 <- ggplot(data=stack(subset(c1045.as.column, subset = BrainMask != 0), 
                        select = c(FALSE,FALSE,FALSE,paradigm$Random) )) +
  geom_histogram(aes(x=values))

p3 <- ggplot(data=stack(subset(c1045.as.column, subset = BrainMask != 0), 
                        select = ! (names(c1045.as.column) %in% c("BrainMask", "ZStat", "ClusterMask")) )) +
  geom_histogram(aes(x=values))

grid.arrange(p1,p2,p3,ncol=3)
```



##Same histograms as above, but ONLY only voxels deemed "active" in biological task, compared to random task
```{r}
#plot histogram of time activity for different tasks, over ONLY active voxels
p1 <- ggplot(data=stack(subset(c1045.as.column, subset = BrainMask != 0 & ClusterMask > 0), 
                        select = c(FALSE,FALSE,FALSE,paradigm$Biological) )) +
  geom_histogram(aes(x=values))

p2 <- ggplot(data=stack(subset(c1045.as.column, subset = BrainMask != 0 & ClusterMask >0 ), 
                        select = c(FALSE,FALSE,FALSE,paradigm$Random) )) +
  geom_histogram(aes(x=values))

p3 <- ggplot(data=stack(subset(c1045.as.column, subset = BrainMask != 0 & ClusterMask > 0), 
                        select = ! (names(c1045.as.column) %in% c("BrainMask", "ZStat", "ClusterMask")) )) +
  geom_histogram(aes(x=values))

grid.arrange(p1,p2,p3,ncol=3)
```

##Scatterplot of BOLD Activity vs Z Statistic (for biological > random) over time points with biological task ON and OFF, and time points with random task ON and OFF
```{r}
#gather all time activitry points into one column, for all masked points.
#introduce group factors for biological vs motion, and activation region vs. not activated
result <- subset(c1045.as.column, subset = BrainMask != 0) %>%
  gather(key = "Label", value = "Activity", 4:ncol(c1045.as.column)) %>%
  mutate(on.off = Label %in% names(c1045.as.column)[c(FALSE, FALSE, FALSE, paradigm$Biological)])
result$color.groups <- "None"
result$color.groups[result$ClusterMask == 0 & result$on.off == TRUE] <- "1 Biological, Not Active"
result$color.groups[result$ClusterMask == 0 & result$on.off == FALSE] <- "2 Random, Not Active"
result$color.groups[result$ClusterMask > 0 & result$on.off == TRUE] <- "3 Biological, Activated"
result$color.groups[result$ClusterMask > 0 & result$on.off == FALSE] <- "4 Random, Activated"

result$color.groups <- as.factor(result$color.groups)

#get just a small sample (1%) of the activity data, otherwise too many voxels
sampleres <- result %>% group_by(on.off) %>% sample_frac(size = .01, replace = FALSE)


#scatterplot of Z Statistic (biological > random) vs activity levels,
#for biological task ON, random taks ON
ggplot() +
  geom_point(data = sampleres, aes(x=ZStat, y=Activity, color=color.groups)) +
  labs(y = "BOLD Intensity", x="Z Statistic") + 
  geom_line(data = sampleres, aes(x=floor(ZStat*10)/10,y=floor(Activity/5)*5, color = paste("Average over activation = ",on.off)), stat="summary",   fun.y = mean) +
  scale_color_manual(values=c("#FFCCFF", "#9999CC", "#66CC99", "#3399CC", "Red", "Blue"))
```

##Same plot as above, but with the BOLD signal at each voxel noramlized (means subtracted from each voxel's time plot)
```{r}
###repeat the above, but using CENTERED fMRI data (mean centered at 0)
c1045.as.column.centered <- c1045.as.column
c1045.as.column.centered[,4:ncol(c1045.as.column.centered)] <-
  sweep( c1045.as.column.centered[,4:ncol(c1045.as.column.centered)], 1, 
         apply(c1045.as.column.centered[,4:ncol(c1045.as.column.centered)], 1, mean),
         "-"  )

result <- subset(c1045.as.column.centered, subset = BrainMask != 0) %>%
  gather(key = "Label", value = "Activity", 4:ncol(c1045.as.column.centered)) %>%
  mutate(on.off = Label %in% names(c1045.as.column.centered)[c(FALSE, FALSE, FALSE, paradigm$Biological)])
result$color.groups <- "None"
result$color.groups[result$ClusterMask == 0 & result$on.off == TRUE] <- "1 Biological, Not Active"
result$color.groups[result$ClusterMask == 0 & result$on.off == FALSE] <- "2 Random, Not Active"
result$color.groups[result$ClusterMask > 0 & result$on.off == TRUE] <- "3 Biological, Activated"
result$color.groups[result$ClusterMask > 0 & result$on.off == FALSE] <- "4 Random, Activated"

result$color.groups <- as.factor(result$color.groups)

#get just a small sample (1%) of the activity data, otherwise too many voxels
sampleres <- result %>% group_by(on.off) %>% sample_frac(size = .01, replace = FALSE)

#scatterplot of Z Statistic (biological > random) vs activity levels,
#for biological task ON, random taks ON
ggplot() +
  geom_point(data = sampleres, aes(x=ZStat, y=Activity, color=color.groups)) +
  labs(y = "BOLD Signal Level", x="Z Stat") + 
  geom_line(data = sampleres, aes(x=floor(ZStat*10)/10,y=floor(Activity/5)*5, color = paste("Average over activation = ",on.off)), stat="summary",   fun.y = mean) +
  scale_color_manual(values=c("#FFCCFF", "#9999CC", "#66CC99", "#3399CC", "Red", "Blue"))
```

###Looking at relationship between BOLD activity levels, between biological, random
##Histograms ONLY over active voxels
```{r}



###Looking at relationship between activity levels, between biological, random
##plot histogram of activity levels for different tasks, over ONLY active voxels
# p1 <- ggplot(data=stack(subset(c1045.as.column.centered, subset = BrainMask != 0 & ClusterMask > 0), 
#                         select = -c(1,2,3) )) +
#   geom_histogram(aes(x=values), binwidth = 3) +
#   scale_x_continuous(limits = c(-100,100))
# 
# temp <- stack(subset(c1045.as.column.centered, subset = BrainMask != 0 & ClusterMask ==0 ), 
#               select = -c(1,2,3) )
# temp <- sample_frac(temp, size=.1, replace=FALSE)
# 
# 
# p2 <- ggplot(data=temp) +
#   geom_histogram(aes(x=values), binwidth = 3) +
#   scale_x_continuous(limits = c(-100,100))
# 
# grid.arrange(p1,p2,ncol=2)


temp <- gather(subset(c1045.as.column.centered, subset = BrainMask != 0),
               key = "TimePoint", value = "result", 4:ncol(c1045.as.column.centered))
temp <- sample_frac(temp, size=.1, replace=FALSE)
ggplot(data=temp) +
  geom_boxplot(aes(x= ClusterMask == 0, y=result)) +
  ylim(c(-200,200)) +
  labs(x="Active Region?", y="BOLD Intensity")
 




##histogram of Z scores
#ggplot(data=sample) +
#  geom_histogram(aes(x=.5*log((1+ZStat)/(1-ZStat)))) +
#  labs(x="Fisher transform of Z Stat") +
#  facet_wrap(~on.off) 

##boxplot of Z scores for on and off
#ggplot(data=sample) +
#  geom_boxplot(aes(x=on.off, y=ZStat))


```



##Histogram of % change between (a statistic of) biological motion time activity points, and (a statistic of) random motion time activity points
```{r}

temp <- as.matrix(subset(c1045.as.column, subset = BrainMask != 0, select = ! names(c1045.as.column) %in% c("BrainMask", "ZStat", "ClusterMask", "corr.biological", "bio.rand.max.min.diff")))

stats.activities.bio <- apply(temp[,paradigm$Biological], 1, 
                              function(x) {
                                c(max = max(x),
                                  median = median(x),
                                  mean = mean(x),
                                  quantile90 = quantile(x,probs=.9),
                                  quantile80 = quantile(x,probs=.8),
                                  quantile70 = quantile(x,probs=.7))       
                                
                              })


temp <- as.matrix(subset(c1045.as.column, subset = BrainMask != 0, select = ! names(c1045.as.column) %in% c("BrainMask", "ZStat", "ClusterMask", "corr.biological", "bio.rand.max.min.diff")))


stats.activities.rand <- apply(temp[,paradigm$Random], 1, 
                              function(x) {
                                c(max = max(x),
                                  median = median(x),
                                  mean = mean(x),
                                  quantile90 = quantile(x,probs=.9),
                                  quantile80 = quantile(x,probs=.8),
                                  quantile70 = quantile(x,probs=.7))       
                                
                              })

temp <- data.frame(matrix(0, nrow = nrow(c1045.as.column), ncol = 6))
temp[c1045.as.column$BrainMask != 0,] <- t(stats.activities.bio - stats.activities.rand)
names(temp) <- c("bio.rand.max.min.diff", "bio.rand.median.diff",
                 "bio.rand.mean.diff", "bio.quantile90.diff",
                 "bio.quantile80.diff", "bio.quantile70.diff")

c1045.as.column <- cbind(c1045.as.column, temp)


rm(list = c(stats.activities.bio, stats.activities.rand))

p1 <- ggplot(data = subset(c1045.as.column, BrainMask != 0), 
       aes(x=ZStat, y=bio.rand.max.min.diff)) +
  geom_point(aes(color=ClusterMask)) +
    geom_line(aes(x=floor(ZStat*10)/10,y=floor(bio.rand.max.min.diff/5)*5),stat = "summary", fun.y = mean) +
  labs(x = "Z Statistic", y = "Max over biotask points MINUS min over randtask points")


p2 <- ggplot(data = subset(c1045.as.column, BrainMask != 0), 
       aes(x=ZStat, y=bio.rand.median.diff)) +
  geom_point(aes(color=ClusterMask)) +
    geom_line(aes(x=floor(ZStat*10)/10,y=floor(bio.rand.median.diff/5)*5),stat = "summary", fun.y = mean) +
    labs(x = "Z Statistic", y = "Median of biotask points MINUS Median of randtask points")



p3 <- ggplot(data = subset(c1045.as.column, BrainMask != 0), 
       aes(x=ZStat, y=bio.rand.mean.diff)) +
  geom_point(aes(color=ClusterMask)) +
    geom_line(aes(x=floor(ZStat*10)/10,y=floor(bio.rand.mean.diff/5)*5),stat = "summary", fun.y = mean) +
    labs(x = "Z Statistic", y = "Mean over biotask points MINUS Mean over randtask points")

grid.arrange(p1,p2,p3,ncol = 3)
```

```{r}
p1<- ggplot(data = subset(c1045.as.column, BrainMask != 0), 
       aes(x=ZStat, y=bio.quantile90.diff)) +
  geom_point(aes(color=ClusterMask)) +
    geom_line(aes(x=floor(ZStat*10)/10,y=floor(bio.quantile90.diff/5)*5),stat = "summary", fun.y = mean) +
    labs(x = "Z Statistic", y = "90% quartile over biotask points MINUS over randtask points")
  


p2<- ggplot(data = subset(c1045.as.column, BrainMask != 0), 
       aes(x=ZStat, y=bio.quantile80.diff)) +
  geom_point(aes(color=ClusterMask)) + 
    geom_line(aes(x=floor(ZStat*10)/10,y=floor(bio.quantile80.diff/5)*5),stat = "summary", fun.y = mean) +
    labs(x = "Z Statistic", y = "80% quartile over biotask points MINUS over randtask points")


p3<- ggplot(data = subset(c1045.as.column, BrainMask != 0), 
       aes(x=ZStat, y=bio.quantile70.diff)) +
  geom_point(aes(color=ClusterMask)) +
    geom_line(aes(x=floor(ZStat*10)/10,y=floor(bio.quantile70.diff/5)*5),stat = "summary", fun.y = mean) +
  labs(x = "Z Statistic", y = "70% quartile over biotask points MINUS over randtask points")


grid.arrange(p1, p2,p3,ncol = 3)

c1045.as.column$bio.rand.max.min.diff <- NULL
c1045.as.column$bio.rand.median.diff <- NULL
c1045.as.column$bio.rand.mean.diff <- NULL
c1045.as.column$bio.rand.quantile90.diff <- NULL
c1045.as.column$bio.rand.quantile80.diff <- NULL
c1045.as.column$bio.rand.quantile70.diff <- NULL

```

##Compute correlation coefficients between biological even paradigm, and time activity
##signal for all voxels

```{r}
#c1045.as.column$corr.biological <- apply(subset(c1045.as.column, 
#                                                select = ! names(c1045.as.column) %in% c("BrainMask", "ZStat", "ClusterMask", "corr.biological", "bio.rand.max.min.diff")),
#                                         1, cor, paradigm$Biological)

c1045.as.column$corr.biological <- apply(c1045.as.column[,c(4:(nrow(paradigm)+3))],
                                         1, cor, paradigm$Biological)

temp <- sample_frac(subset(c1045.as.column, BrainMask != 0), size = .1, replace = FALSE)

ggplot(data = temp, aes(x=ZStat, y=.5*log((1+corr.biological)/(1-corr.biological)))) +
  geom_point(aes(color=ClusterMask)) +
  labs(y="Fisher transform of Correlation Coefficient")

```



