---
title: "Modeling Portion of Capstone Project"
author: "Michael An"
date: "April 19, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caTools)
library(ROCR)
library(glmnet)
library(caret)
```


```{r, echo=FALSE}
source("/home/mike/Documents/Foundations_Course/Capstone/exploratory_script_load.R")

#Attach computed statistics to c1045.as.column.centered variable as well
#In the process, have to recompute some statistics from EDA section 
#for 2 run data

temp <- as.matrix(subset(c1045.as.column, subset = BrainMask != 0, select = ! names(c1045.as.column) %in% c("BrainMask", "ZStat", "ClusterMask", "corr.biological", "bio.rand.max.min.diff")))

stats.activities.bio <- apply(temp[,paradigm$Biological], 1, 
                              function(x) {
                                c(max = max(x),
                                  median = median(x),
                                  mean = mean(x),
                                  quantile90 = quantile(x,probs=.9),
                                  quantile80 = quantile(x,probs=.8),
                                  quantile70 = quantile(x,probs=.7))       
                                
                              })


temp <- as.matrix(subset(c1045.as.column, subset = BrainMask != 0, select = ! names(c1045.as.column) %in% c("BrainMask", "ZStat", "ClusterMask", "corr.biological", "bio.rand.max.min.diff")))


stats.activities.rand <- apply(temp[,paradigm$Random], 1, 
                              function(x) {
                                c(max = max(x),
                                  median = median(x),
                                  mean = mean(x),
                                  quantile90 = quantile(x,probs=.9),
                                  quantile80 = quantile(x,probs=.8),
                                  quantile70 = quantile(x,probs=.7))       
                                
                              })

temp <- data.frame(matrix(0, nrow = nrow(c1045.as.column), ncol = 6))
temp[c1045.as.column$BrainMask != 0,] <- t(stats.activities.bio - stats.activities.rand)
names(temp) <- c("bio.rand.max.min.diff", "bio.rand.median.diff",
                 "bio.rand.mean.diff", "bio.quantile90.diff",
                 "bio.quantile80.diff", "bio.quantile70.diff")

c1045.as.column <- cbind(c1045.as.column, temp)

c1045.as.column$corr.biological <- apply(c1045.as.column[,c(4:(nrow(paradigm)+3))],
                                         1, cor, paradigm$Biological)


c1045.as.column.centered$bio.quantile90.diff <- 
  c1045.as.column$bio.quantile90.diff
c1045.as.column.centered$bio.quantile80.diff <-
  c1045.as.column$bio.quantile80.diff
c1045.as.column.centered$bio.quantile70.diff <-
  c1045.as.column$bio.quantile70.diff
c1045.as.column.centered$corr.biological <- 
  c1045.as.column$corr.biological



temp <- as.matrix(subset(c1045.as.column.centered.2runs, subset = BrainMask != 0, select = ! names(c1045.as.column.centered.2runs) %in% c("BrainMask", "ZStat", "ClusterMask", "corr.biological", "bio.rand.max.min.diff")))

stats.activities.bio <- apply(temp[,paradigm.2runs$Biological], 1, 
                              function(x) {
                                c(max = max(x),
                                  median = median(x),
                                  mean = mean(x),
                                  quantile90 = quantile(x,probs=.9),
                                  quantile80 = quantile(x,probs=.8),
                                  quantile70 = quantile(x,probs=.7))       
                                
                              })


temp <- as.matrix(subset(c1045.as.column.centered.2runs, subset = BrainMask != 0, select = ! names(c1045.as.column.centered.2runs) %in% c("BrainMask", "ZStat", "ClusterMask", "corr.biological", "bio.rand.max.min.diff")))


stats.activities.rand <- apply(temp[,paradigm.2runs$Random], 1, 
                              function(x) {
                                c(max = max(x),
                                  median = median(x),
                                  mean = mean(x),
                                  quantile90 = quantile(x,probs=.9),
                                  quantile80 = quantile(x,probs=.8),
                                  quantile70 = quantile(x,probs=.7))       
                                
                              })

temp <- data.frame(matrix(0, nrow = nrow(c1045.as.column.centered.2runs), ncol = 6))
temp[c1045.as.column.centered.2runs$BrainMask != 0,] <- t(stats.activities.bio - stats.activities.rand)
names(temp) <- c("bio.rand.max.min.diff", "bio.rand.median.diff",
                 "bio.rand.mean.diff", "bio.quantile90.diff",
                 "bio.quantile80.diff", "bio.quantile70.diff")

c1045.as.column.centered.2runs <- cbind(c1045.as.column.centered.2runs, temp)

c1045.as.column.centered.2runs$corr.biological <- apply(c1045.as.column.centered.2runs[,c(4:(nrow(paradigm.2runs)+3))],
                                         1, cor, paradigm.2runs$Biological)
```

###Fitting a simple logistic model
```{r, echo=FALSE}
#Set generic names for the data, so you dont change the originals
subject.image.mask <- c1045.mask
model.data <- subset(c1045.as.column.centered, BrainMask != 0)
model.data$ClusterMaskBinary <- model.data$ClusterMask > 0
model.paradigm <- paradigm



data.split.bool <- sample.split(model.data$ClusterMaskBinary, SplitRatio = .25)

data.train <- subset(model.data, data.split.bool = TRUE)
data.test <- subset(model.data, data.split.bool = FALSE)
```

```{r}
#Logistic regression model
ActivationModel <- glm(ClusterMaskBinary ~ . - BrainMask - ZStat - ClusterMask - ClusterMaskBinary , data = data.train, family = binomial)

predictionTrain <- predict(ActivationModel, type="response" )

tapply(predictionTrain, data.train$ClusterMaskBinary, mean)

summary(ActivationModel)
```


#Drop the correlation coefficient as a covaraite, and rerun the logistic model fit

```{r}
#Fit logistic model over data with corr.biological removed
ActivationModel <- glm(ClusterMaskBinary ~ . - BrainMask - ZStat - ClusterMask - ClusterMaskBinary - corr.biological, data = data.train, family = binomial)

predictionTrain <- predict(ActivationModel, type="response" )

tapply(predictionTrain, data.train$ClusterMaskBinary, mean)

summary(ActivationModel)
```

#assess z stats of coefficients corresponding to time activity paradigm
```{r, echo=FALSE}

model.paradigm$pval <- coef(summary(ActivationModel))[,4][2:(nrow(model.paradigm)+4-3)] < 5e-2
```



#Analyze prediction accuracy using ROC curve and determine threshold to minimize false pos rate
```{r}

#Use ROCR library to plot ROC curve
cluster.pred <- prediction(predictionTrain, model.data$ClusterMaskBinary)
perform.func <- performance(cluster.pred, "tpr", "fpr")

plot(perform.func, print.cutoffs.at=seq(.1,.9,.1))
```

```{r}

#verify model by predicting the combined training and test set
predictionTest <- predict(ActivationModel, type="response", newdata = model.data)

confuse.matrix <- table(model.data$ClusterMaskBinary,
                        predictionTest > .3)

#sensitivity - true positive rate
confuse.matrix[2,2]/(confuse.matrix[2,1]+confuse.matrix[2,2])

#specificity - true negative rate
confuse.matrix[1,1]/(confuse.matrix[1,1]+confuse.matrix[1,2])
```



#overlay activation map onto a reference image, and compare with precomputed group map 
```{r, echo=FALSE}
#create test cluster image based on the given prediction threshold

prediction.all <- predict(ActivationModel, type="response", 
                          newdata = model.data)

#reference Nifti object to put out new mask into
nifti.prediction.mask <- c1045.cluster.mask

#create new space to put in the mask in image space
#using the BrainMask > 0 mask
temp <- array(0, dim = dim(nifti.prediction.mask@.Data))
temp[subject.image.mask@.Data > 0] <- as.integer( prediction.all > .3 )

#insert the data into the nifti object's .Data field 
nifti.prediction.mask@.Data <- temp
```

```{r}
#write and save resulting maps

#writeNIfTI(nifti.prediction.mask, "c1045_logistic_classific_map", onefile=TRUE)

#View the resulting image and mask
overlay(c1045, ifelse(nifti.prediction.mask == 0, NA, nifti.prediction.mask), col.y = heat.colors(1))


```

```{r}
overlay(c1045, ifelse(c1045.cluster.mask ==0, NA, 1),col.y=heat.colors(1))

```


#Repeat analysis above, but with using only 2 run's time points as predictors
```{r, echo=FALSE}
#Set generic names for the data, so you dont change the originals
model.data <- subset(c1045.as.column.centered.2runs, BrainMask != 0)
model.data$ClusterMaskBinary <- model.data$ClusterMask > 0
model.paradigm <- paradigm.2runs

data.split.bool <- sample.split(model.data$ClusterMaskBinary, SplitRatio = .25)

data.train <- subset(model.data, data.split.bool = TRUE)
data.test <- subset(model.data, data.split.bool = FALSE)
```

#Model Summary and Analysis for 2 run version
```{r}
#fit logistic model over 2 run worth of data
ActivationModel <- glm(ClusterMaskBinary ~ . - BrainMask - ZStat - ClusterMask - ClusterMaskBinary - corr.biological, data = data.train, family = binomial)

predictionTrain <- predict(ActivationModel, type="response" )

summary(ActivationModel)
```
```{r}

#Use ROCR library to plot ROC curve
cluster.pred <- prediction(predictionTrain, model.data$ClusterMaskBinary)
perform.func <- performance(cluster.pred, "tpr", "fpr")

plot(perform.func, print.cutoffs.at=seq(.1,.9,.1))
```
```{r}

#verify model by predicting the combined training and test set
predictionTest <- predict(ActivationModel, type="response", newdata = model.data)

confuse.matrix <- table(model.data$ClusterMaskBinary,
                        predictionTest > .3)

#sensitivity - true positive rate
confuse.matrix[2,2]/(confuse.matrix[2,1]+confuse.matrix[2,2])

#specificity - true negative rate
confuse.matrix[1,1]/(confuse.matrix[1,1]+confuse.matrix[1,2])
```

#Generate images for 2 run results and compare to group map
```{r, echo=FALSE}
#create test cluster image based on the given prediction threshold

prediction.all <- predict(ActivationModel, type="response", 
                          newdata = model.data)

#reference Nifti object to put out new mask into
nifti.prediction.mask <- c1045.cluster.mask

#create new space to put in the mask in image space
#using the BrainMask > 0 mask
temp <- array(0, dim = dim(nifti.prediction.mask@.Data))
temp[subject.image.mask@.Data > 0] <- as.integer( prediction.all > .3 )

#insert the data into the nifti object's .Data field 
nifti.prediction.mask@.Data <- temp
```
```{r}
#write and save resulting maps

#writeNIfTI(nifti.prediction.mask, "c1045_logistic_classific_map", onefile=TRUE)

#View the resulting image and mask
overlay(c1045, ifelse(nifti.prediction.mask == 0, NA, nifti.prediction.mask), col.y = heat.colors(1))


```
```{r}
overlay(c1045, ifelse(c1045.cluster.mask ==0, NA, 1),col.y=heat.colors(1))

```

#Explore PCA on the fMRI time series for the 2 run subset, and trian logistic model on PCA components instead of normal time points
```{r}
#generate PCA object from centered 2 run data
princomp.obj <- preProcess(subset(c1045.as.column.centered.2runs,
                          select = names(c1045.as.column.centered.2runs) %in% 
                            grep("X", names(c1045.as.column.centered.2runs), value = TRUE)), 
                   method=c("BoxCox", "center", 
                            "scale", "pca") )

#get the principal components
PC <- predict(princomp.obj, subset(c1045.as.column.centered.2runs,
                          select = names(c1045.as.column.centered.2runs) %in% 
                            grep("X", names(c1045.as.column.centered.2runs), value = TRUE)))


PC$ClusterMaskBinary <- c1045.as.column.centered.2runs$ClusterMask > 0

data.split.bool <- sample.split(PC$ClusterMaskBinary, SplitRatio = .25)

data.train <- subset(PC, data.split.bool == TRUE)
data.test <- subset(PC, data.split.bool == FALSE)

#fit a logistic model using the PCs found
ActivationModel <- glm(ClusterMaskBinary ~ ., family=binomial, data =PC[,-(29:(ncol(PC)-1))] )
prediction.all <- predict(ActivationModel, type="response")


#ROC curve of pred results
cluster.pred <- prediction(prediction.all, PC$ClusterMaskBinary)
perform.func <- performance(cluster.pred, "tpr", "fpr")

plot(perform.func, print.cutoffs.at=seq(.1,.9,.1))
```

```{r}
#confusion matrix
confuse.matrix <- table(PC$ClusterMaskBinary, prediction.all > .2)
confuse.matrix

confuse.matrix[2,2]/(confuse.matrix[2,1]+confuse.matrix[2,2])

confuse.matrix[1,1]/(confuse.matrix[1,1]+confuse.matrix[1,2])

```

